Nonlinear Spatio-Temporal Modelling using the Integro-Difference Equation

### 2010-07-08 ltx/second_order_framework/second_order_manuscript.tex
some small edits to 2nd order paper and fixed typo in DataPaper data viewing script

%to delete
\newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} 
%opening
\title{Non-linear Spatio-temporal modelling using the Integro-Difference Equation}
\author{Parham Aram*,
        Dean~R.~Freestone,Kenneth Scerri, Michael Dewar,\\ Andrew Zammit Mangion and Visakan Kadirkamanathan}

@@

\section{Introduction}
\section{State space Representation of the Spatio-temporal Non-linear IDE}
We define the $p$th order non-linear, stochastic, spatio-temporal homogenous IDE as
 \begin{equation}\label{eq:IDEModel}
  \rho_t(\mathbf r)=\int_{\Omega}w(\mathbf r-\mathbf r')f(v_{t-p}(\mathbf r'))d\mathbf r'+\mathbf e_{t-p}(\mathbf r) 
 \end{equation}
where $\rho_t(\mathbf r)$ is a lag polynomial defined as
\begin{equation}\label{eq:LagPolynomial}
 \rho_t(\mathbf r)=(1+\sum_{i=1}^{p}\varphi_i L_i)v_{t}\left(\mathbf r\right)
\end{equation}

where $v_t(\mathbf r)$ denotes the spatial field at time $t$ and at spatial location $\mathbf r$, $f(.)$ is some nonlinear map, and $w(.)$ is the spatial kernel. The disturbance $\mathbf e_{t-p}(\mathbf r)$ is a zero-mean normally distributed noise process, spatially
coloured but temporally independent, with covariance

@@

where $\delta(.)$ is the Dirac-delta function.
 The spatial field is observed by

@@

where the observation vector $\mathbf{y}_{t}$, compiled at $n_{y}$ spatial locations and at time $t$, is corrupted by an i.i.d normally distributed zero-mean white noise ${\boldsymbol\varepsilon}_t\sim \mathcal{N}\left(\mathbf{0},\boldsymbol\Sigma_{\varepsilon}\right)$ with $\mathbf{\Sigma}_{\boldsymbol\varepsilon}=\sigma_{\varepsilon}^2\mathbf I_{n_y} $. It is assumed here that $\boldsymbol\Sigma_{\varepsilon}$ and $\gamma\left(\mathbf r-\mathbf r' \right)$ are known. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry and is defined by 

@@

where $\sigma_m$ sets the sensor width. The superscript $\top$ denotes the transpose operator. We proceed by writing the spatio-temporal IDE model in a parameterized state space form which allows the application of standard estimation
techniques. We use a decomposition of the filed and the spatial kernel using a set of Gaussian basis functions

@@

\subsection*{Discritization}

@@

\subsection*{From Difference equation to state space model}

@@

dv(\mathbf r,t)=\tilde{v}(\mathbf r,t)d_t & \\

>>>>

%to delete
\newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} 
%opening
\title{Nonlinear Spatio-Temporal Modelling using the Integro-Difference Equation}
\author{Parham Aram*,
        Dean~R.~Freestone,Kenneth Scerri, Michael Dewar,\\ Andrew Zammit Mangion and Visakan Kadirkamanathan}

@@

\section{Introduction}
\section{State-Space Representation of the Spatio-temporal Nonlinear IDE}
The $p^{th}$-order nonlinear, stochastic, spatio-temporal homogenous IDE model is defined as
 \begin{equation}\label{eq:IDEModel}
  \rho_t(\mathbf r)=\int_{\Omega}w(\mathbf r-\mathbf r')f(v_{t-p}(\mathbf r'))d\mathbf r'+\mathbf e_{t-p}(\mathbf r). 
The quantity $\rho_t(\mathbf r)$ is a lag polynomial defined as
\begin{equation}\label{eq:LagPolynomial}
 \rho_t(\mathbf r)=(1+\sum_{i=1}^{p}\varphi_i L_i)v_{t}\left(\mathbf r\right),
where $v_t(\mathbf r)$ denotes a spatio-temporal field at time $t$ and at spatial location $\mathbf r$, $f(.)$ is a nonlinear map, and $w(.)$ is a spatial mixing kernel. The disturbance, $\mathbf e_{t-p}(\mathbf r)$, is a zero-mean normally distributed noise process that is spatially
coloured and temporally white, with covariance

@@

and $\delta(.)$ is the Dirac-delta function.
Observations from the spatio-temporal field are described by the measurement function

@@

where the observation vector, $\mathbf{y}_{t}$, compiled at $n_{y}$ spatial locations and at time $t$, is corrupted by an i.i.d normally distributed zero-mean white noise ${\boldsymbol\varepsilon}_t\sim \mathcal{N}\left(\mathbf{0},\boldsymbol\Sigma_{\varepsilon}\right)$ with $\mathbf{\Sigma}_{\boldsymbol\varepsilon}=\sigma_{\varepsilon}^2\mathbf I_{n_y} $. It is assumed here that $\boldsymbol\Sigma_{\varepsilon}$ and $\gamma\left(\mathbf r-\mathbf r' \right)$ are known. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry and is defined by 

@@

where $\sigma_m$ sets the sensor width. The superscript $\top$ denotes the transpose operator. We proceed by writing the spatio-temporal IDE model in a parameterized state space form which allows the application of standard estimation techniques. We use a decomposition of the field and the spatial mixing kernel using a set of Gaussian basis functions

@@

\subsection*{Discretization}

@@

\subsection*{From Difference Equation to State-Space Model}

@@

dv(\mathbf r,t)=\tilde{v}(\mathbf r,t)dt & \\
